The symmetry of partner modelling

### 2016-02-04 mutual-modeling.tex
Updated section 2

\section{Mutual Modelling}

We refer to \emph{mutual modelling} as the process of inferring one's partner
mental states. Any claim that students carry out a detailed monitoring of their
peers would be as incorrect as any claim that they do not maintain any
representation at all. If mutual modelling had to be permanently detailed and
accurate, subjects would obviously face a huge cognitive load. Conversely, peers
could not collaborate without some minimal amount of mutual modelling. For
instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
profession) and, of course, it can be incorrect\ldots but it can not remain empty.
Dialogues include many instances of utterances such as ``I thought he would do
that'' (first level of mutual modelling) or even ``He thought I would do that
but I intended something else.'' (second level of mutual modelling).

The content of mutual models ranges from \emph{dispositional} to
\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
representation of $B$'s long term knowledge, skills or traits. It is thus closely
related to the notion of transactive memory~\citep{wegner1987transactive,
moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
B's knowledge, behavior or intentions specifically activated in the situation in
which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
ones for 2 hours. Examples of fragments that constitute $A$'s model of B
regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:

@@

The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
beliefs, emotions, history, status, etc.

Besides, we have different levels of mutual modelling. If $A$ states ``B thinks I am good in
maths'', $A$ builds a \emph{second level} model:
\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
regress of nested models: $A$
saying ``$B$ knows that I don't expect him to solve this statistics problem''
corresponds to \\ \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.

A partner model is likely not a ``box'', \ie not a monolithic representation
but rather a mosaic of information fragments about the partner, with various
granularity and various life cycles. This mosaic is elaborated through a variety
of mechanisms, first for building an initial model of the partner, then for
updating this model. As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other based on cues such as
his/her belonging to broad categories (age, culture, profession, \ldots),
stereotypes (sportsmen, junkie, business women, Swiss,\ldots) as well as physical
appearance. Scholars studied how initial modelling impacts communication. In
their experiments on initial mutual modelling, \citet{slugoski1993attribution}
pretended to their subjects that their (fake) partner had or had not received
the same information. They observed that the subjects adapted their dialogue by
focusing the explanation on the items that he/she was supposed to ignore.
\citet{brennan1991conversation} showed that the subjects used different initial
strategies in forming queries depending on who they were told their partner was.

@@

After initialization, mutual models are updated during the collaborative work
through verbal and non-verbal interactions. A default inference rule is that ``my
partner agrees with me unless he disagrees'', which rejects the critiques that
mutual modelling generates an unbearable cognitive load. This default rule is
superseded by the several mechanisms for monitoring and repairing the partner
understanding: acknowledgement, continuous attention, relevance of next turns,
facial expressions including gaze signals, etc.

Mutual modelling does not occur in a vacuum but it is highly contextualized.
\citet{clark1991grounding} review how the features of the collaborative
situation, namely the media (co-temporality,\ldots), may facilitate or hamper
mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
short silence between two pilots was perfectly interpreted because it occurred
in a highly constrained communication context.  Some environments are more
productive than others in helping peers to detect their misunderstandings.
\citet{roschelle1995construction} reformulate the design of CSCL
interfaces to provide ways for peers to detect and repair their
misunderstanding. In CSCW, researchers have explored various so-called
``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
that A could not directly perceive because $B$ was working on a different subset
of the virtual space. Different awareness tools are presented and relied upon in
this contribution as well since they allowed us to manipulate experimentally the
mutual modelling activity.

%\subsection{State of the Art}
%
%\paragraph{Building a model of a human during the interaction} Perspective
%Taking is a human ability which allows one to put him/herself in another
%person's point of view. Studied in psychology
%literature~\citep{Flavell1992,Tversky1999}, this ability is crucial when
%interacting with people by allowing one to reason on others' understanding of
%the world in terms of visual perception, spatial descriptions, affordances and
%beliefs, etc.  In the last years these notions have been gradually employed in
%HRI.~\citep{Breazeal2006} presents a learning algorithm that takes into account
%information about a teacher's visual perspective in order to learn a task.
%~\citep{Johnson2005} apply visual perspective taking for action recognition
%between two robots.~\citep{Trafton2005} use both visual and spatial perspective
%taking to find out the referent indicated by a human partner.
%
%The cognitive model that the robot builds for the agent it interacts with is
%still simple and mostly focused on geometric features (who sees what? What are
%our relative positions? etc.). Extending this knowledge with more subtle
%perceptions (emotional state for instance) remains to be done.
%
%\paragraph{Joint action}
%Several theories dealing with collaboration~\citep{Cohen1991,Grosz1996,Clark1996}
%emphasize that collaborative tasks have specific requirements compared to
%individual ones, \eg, since the robot and the person share a common goal, they
%have to agree on the manner to realize it, they must show their commitment to
%the goal during execution, etc. Several robotic systems have already been built
%based on these theories~\citep{Rich1997,Sidner2005,Breazeal2003} and they all
%have shown benefits of this approach. They have also shown how difficult it is
%to manage turn-taking between communication partners and to interleave task
%realization and communication in a generic way. Finally, today only few
%systems~\citep{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
%levels.
%

>>>>

\section{Partner Modelling}

We refer to \emph{Partner modelling} as the process of inferring one's partner
mental states. Any claim that students carry out a detailed monitoring of their
peers would be as incorrect as any claim that they do not maintain any
representation at all. If mutual modelling had to be permanently detailed and
accurate, subjects would obviously face a huge cognitive load. Conversely, peers
could not collaborate without some minimal amount of mutual modelling. 
Collaborative learning dialogue include many instances of utterances such as ``I thought he would do
that'' (first level of mutual modelling) or even ``He thought I would do that
but I intended something else.'' (second level of mutual modelling).

%For instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
%opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
%B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
%evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
%it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
%profession) and, of course, it can be incorrect\ldots but it can not remain empty.

The content of the partner model ranges from \emph{dispositional} to
\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
representation of $B$'s long term knowledge, skills or traits. It is thus closely
related to the notion of transactive memory~\citep{wegner1987transactive,
moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
B's knowledge, behavior or intentions specifically activated in the situation in
which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
ones for 2 hours. Examples of fragments that constitute $A$'s model of B
regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:

@@

The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
beliefs, emotions, history, status, etc. A partner model is likely not a ``box'', \ie not a monolithic representation
but rather a mosaic of information fragments about the partner, with various
granularity and various life cycles. This mosaic is elaborated through a variety
of mechanisms, first for building an initial model of the partner, then for
updating this model. As two students meet for the first time, partners models are
initialized by the assumptions they make upon each other based on cues such as
his/her belonging to broad categories (age, culture, profession, \ldots),
stereotypes (sportsmen, junkie, business women, Swiss,\ldots) as well as physical
appearance. Scholars studied how initial modelling impacts communication. In
their experiment, \citet{slugoski1993attribution}
pretended to their subjects that their (fake) partner had or had not received
the same information. They observed that the subjects adapted their dialogue by
focusing the explanation on the items that he/she was supposed to ignore.
\citet{brennan1991conversation} showed that the subjects used different initial
strategies in forming queries depending on who they were told their partner was.

@@

After initialization, partners models are updated during the collaborative work
through verbal and non-verbal interactions. A default inference rule is that ``my
partner agrees with me unless he disagrees'', which rejects the critiques that
partner modelling generates an unbearable cognitive load. This default rule is
superseded by the several mechanisms for monitoring and repairing the partner
understanding: acknowledgement, continuous attention, relevance of next turns,
facial expressions including gaze signals, etc.

Finally, partner modelling does not occur in a vacuum but it is highly contextualized.
\citet{clark1991grounding} review how the features of the collaborative
situation, namely the media (co-temporality,\ldots), may facilitate or hamper
mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
short silence between two pilots was perfectly interpreted because it occurred
in a highly constrained communication context.  Some environments are more
productive than others in helping peers to detect their misunderstandings.
\citet{roschelle1995construction} reformulate the design of CSCL
interfaces to provide ways for peers to detect and repair their
misunderstanding.
